New Bayes covariant coupling methods to cope with change of priors in one-vs-one classification

### 2025-07-15 intro.tex
Include proper prerequisities;

Distribution shift is an important problem facing machine-learning practitioners  [\cite{zhang2023dive, sugiyama2007covariate}].  Using probabilistic formulation, we say the distribution shift occurs if there is a change  in the joint probability distribution $p(X,Y)$ of covariates $X$ with the dependent variable $Y$  between training and testing. Noting the factorizations 
$$
p(X,Y) = p(Y\mid X)p(X) = p(X \mid Y )p(Y)
$$
one may distinguish the following subcategories of distribution shift:
\begin{itemize}
	\item covariate shift, if $p(X)$ changes, but $p(Y\mid X)$ is unchanged,
	\item label shift, if $p(Y)$ changes, but $P(X \mid Y)$ remains the same,
	\item concept shift, if $p(Y \mid X)$ itself changes.
\end{itemize}

We will be concerned with label shift in classification problems, where $p(Y)$ is just the prior on classes.   To provide a concrete example, consider a driver assistance system recognizing traffic signs. Suppose a speed limit sign was maliciously altered to indicate 80mph limit instead of 30mph. To avoid being fooled by the alteration, the system could utilize priors. The distribution of traffic signs will change markedly when the car leaves highway and enters streets in a city. It is very unlikely to encounter 80mph limit in a city setting. By taking into account the city priors, the system is likely to make the proper classification. In general a well-designed classification system will be able to improve its performance when supplied with a new set of class priors that correspond to the true distribution of classes.

Change of priors can be effectively dealt with in probabilistic classifiers. The basic idea is that Bayes theorem stipulates \emph{exactly} what should happen  to the prediction of the  Bayes classifier when priors change (see \eqref{eq:changePrior}).  In practice we assume that a trained classifier closely approximates the Bayes classifier. Therefore we can make adjustment to our prediction by applying the same transformation of posterior as would happen with the Bayes classifier.

@@

Some aspects of label shift were studied in earlier works including test time adaptation
\cite{vsipka2022hitchhiker}, shift detection and quantification \cite{lipton2018detecting}, statistical properties of label shift estimators \cite{garg2020unified}, and Fisher consistency of estimators of priors \cite{tasche2017fisher}.

Our approach to the problem is to evaluate practical performance of different methods to aggregate binary predictions. Such methods -- called coupling methods -- first appeared in work of \cite{refregier1991probabilistic}. Many authors followed with alternative proposals \cite{price1994pairwise, hastie1998classification,  zahorian1999partitioned, wu2004probability, vsuch2015new, vsuch2016bayes}.

>>>>

Distribution shift is an important problem facing machine-learning practitioners 
\cite{zhang2023dive, sugiyama2007covariate}.  Using probabilistic formulation, we say the distribution shift occurs if there is a change  in the joint probability distribution $p(X,Y)$ of covariates $X$ with the dependent variable $Y$  between training and testing. Noting the factorizations 
$$
p(X,Y) = p(Y\mid X)p(X) = p(X \mid Y )p(Y)
$$
one may distinguish the following subcategories of distribution shift:
\begin{itemize}
	\item covariate shift, if $p(X)$ changes, but $p(Y\mid X)$ is unchanged,
	\item label shift, if $p(Y)$ changes, but $P(X \mid Y)$ remains the same,
	\item concept shift, if $p(Y \mid X)$ itself changes.
\end{itemize}

We will be concerned with label shift in classification problems, where $p(Y)$ 
is just the prior on classes.  Some aspects of label shift were studied in
earlier works including test time adaptation
\cite{vsipka2022hitchhiker}, shift detection and quantification 
\cite{lipton2018detecting}, 
statistical properties of label shift estimators \cite{garg2020unified}, 
and Fisher consistency of estimators of priors \cite{tasche2017fisher}. 
Let us consider  the question 
how to adapt a classifier when the  prior is updated.

To provide a concrete example, consider a driver assistance system recognizing traffic signs. 
Suppose a speed limit sign was maliciously altered to indicate 80mph limit instead of 30mph.
To avoid being fooled by the alteration, the system could utilize priors.
The distribution of traffic signs will change markedly when the car leaves highway and enters streets in a city.
unlikely to encounter 80mph limit in a city setting. By taking into account 
the city priors, the system would be more likely to make the proper classification.

Some systems are unable to do incorporate new priors without rebuilding 
the classifier  e.g. those based 
on fuzzy logic, or some tree methods. 
However, change of priors can be effectively dealt with in probabilistic 
classifiers. The basic idea is that Bayes theorem stipulates \emph{exactly} 
what should happen  to the prediction of the  Bayes classifier when priors 
change (see \eqref{eq:changePrior}).  In practice we assume that a trained 
classifier closely approximates the Bayes classifier. Therefore we can make 
adjustment to our prediction by applying the same transformation of posterior 
as would happen with the Bayes classifier.

@@

Our approach to the problem is to evaluate practical performance of different 
methods to aggregate binary predictions. Such methods -- called coupling methods
-- first appeared in work of \cite{refregier1991probabilistic} when making 
multi-class classification using neural networks. 
Many authors followed with alternative proposals 
\cite{price1994pairwise, hastie1998classification,  zahorian1999partitioned,
wu2004probability, vsuch2015new, vsuch2016bayes}.
